---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
basic_exists <- file.exists(Sys.getenv("EPIEXTRACTS_CPSBASIC_DIR"))
org_exists <- file.exists(Sys.getenv("EPIEXTRACTS_CPSORG_DIR"))
```

# epiextractr

<!-- badges: start -->
<!-- badges: end -->

epiextractr makes it easy to use the [EPI microdata extracts](https://microdata.epi.org/) in R.

## Example
Load a selection of variables from the 2018-2019 EPI CPS ORG extracts:
```{r, eval = org_exists}
library(epiextractr)
load_org(2018:2019, year, female, wage, orgwgt)
```
```{r, echo = FALSE, eval = !org_exists}
cat("
# A tibble: 591,270 x 4
    year orgwgt     female  wage
   <int>  <dbl>  <int+lbl> <dbl>
 1  2018  8544. 0 [Male]    NA  
 2  2018  5985. 1 [Female]  NA  
 3  2018  7412. 1 [Female]  20.8
 4  2018 16209. 1 [Female]  10  
 5  2018  6954. 1 [Female]  25  
 6  2018 10195. 1 [Female]   9.5
 7  2018  8260. 0 [Male]    17  
 8  2018  8544. 0 [Male]    NA  
 9  2018  5985. 1 [Female]  NA  
10  2018  4852. 1 [Female]  NA  
# … with 591,260 more rows
")
```

## Installation and basic usage

First, install the current version of the package from GitHub via devtools 
```{r, eval = FALSE}
devtools::install_github("economic/epiextractr")
```

Then download the CPS microdata using `download_cps()`. For example,
```{r, eval = FALSE}
download_cps("org", "C:\data\cps")
```
will download the latest EPI CPS ORG extracts from https://microdata.epi.org and place them in the directory `C:\data\cps`.

After the data is downloaded, load a selection of CPS data for your analysis:
```{r, eval = FALSE}
load_cps("org", 2000:2019, year, orgwgt, wage, wbho, .extracts_dir = "C:\data\cps")
```

See `vignette("epiextractr")` for more examples.

