---
title: "epiextractr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{epiextractr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

basic_exists <- file.exists(Sys.getenv("EPIEXTRACTS_CPSBASIC_DIR"))
org_exists <- file.exists(Sys.getenv("EPIEXTRACTS_CPSORG_DIR"))

library(epiextractr)
```

The `epiextractr` package makes it easy to use the [EPI microdata extracts](https://microdata.epi.org/) in R. 

First download the CPS microdata using `download_cps()` into a directory on your machine. The downloaded data will simply be .feather files for each year of data. 

For example, to download the Outgoing Rotation Group extracts into a directory called `C:/data/cps`, try
```{r, eval = FALSE}
library(epiextractr)
download_cps("org", "C:/data/cps")
```

Then you can use `load_org()` to call a selection of years and columns:
```{r, eval = FALSE}
load_org(2019:2021, year, female, wage, orgwgt, .extracts_dir = "C:/data/cps")
```
```{r, echo = FALSE, eval = org_exists}
load_org(2019:2021, year, female, wage, orgwgt)
```
```{r, echo = FALSE, eval = !org_exists}
cat("
# A tibble: 824,963 × 4
    year orgwgt female      wage
   <int>  <dbl> <int+lbl>  <dbl>
 1  2019 11367. 1 [Female] 14   
 2  2019  6541. 1 [Female] 20.9 
 3  2019  6327. 0 [Male]    7.65
 4  2019  6327. 0 [Male]    7.65
 5  2019 11262. 1 [Female] 10   
 6  2019  7867. 1 [Female] 28.8 
 7  2019 11262. 1 [Female] 11   
 8  2019  7943. 0 [Male]   NA   
 9  2019  6092. 1 [Female] NA   
10  2019  7738. 0 [Male]   NA   
# … with 824,953 more rows
")
```

For ease of use you can omit the `.extracts_dir` option by setting the following environment variables equal to the directory of the extracts:

```
EPIEXTRACTS_CPSBASIC_DIR
EPIEXTRACTS_CPSORG_DIR
EPIEXTRACTS_CPSMAY_DIR
```

For example, if your .Renviron file sets
```
EPIEXTRACTS_CPSORG_DIR=C:/data/cps
```
then you can simply run
```{r, eval=FALSE}
load_org(2019:2021, year, female, wage, orgwgt)
```

### More examples
#### Calculating employment rates by year and race
Calculate annual employment-to-population ratios by race/ethnicity from the 2010-2019 Basic CPS using tidyverse functions:
```{r, eval = FALSE}
library(tidyverse)
load_cps("basic", 2010:2019, year, basicwgt, wbhao, emp) %>%
  filter(basicwgt > 0) %>%
  group_by(year, wbhao) %>%
  summarize(value = weighted.mean(emp, w = basicwgt)) %>%
  mutate(name = as.character(haven::as_factor(wbhao))) %>%
  pivot_wider(year)
```
```{r, echo=FALSE, message = FALSE, eval = basic_exists}
library(tidyverse)
load_cps("basic", 2010:2019, year, basicwgt, wbhao, emp) %>%
  filter(basicwgt > 0) %>%
  group_by(year, wbhao) %>%
  summarize(value = weighted.mean(emp, w = basicwgt)) %>%
  mutate(name = as.character(haven::as_factor(wbhao))) %>%
  pivot_wider(year)
```
```{r, echo=FALSE, eval = !basic_exists}
cat("
# A tibble: 10 × 6
# Groups:   year [10]
    year White Black Hispanic Asian Other
   <int> <dbl> <dbl>    <dbl> <dbl> <dbl>
 1  2010 0.595 0.523    0.590 0.601 0.501
 2  2011 0.595 0.516    0.589 0.602 0.504
 3  2012 0.594 0.527    0.595 0.604 0.521
 4  2013 0.592 0.529    0.600 0.611 0.499
 5  2014 0.594 0.540    0.612 0.607 0.519
 6  2015 0.596 0.554    0.616 0.606 0.517
 7  2016 0.598 0.562    0.620 0.612 0.525
 8  2017 0.599 0.575    0.627 0.619 0.540
 9  2018 0.601 0.583    0.632 0.619 0.538
10  2019 0.603 0.588    0.639 0.625 0.547
")
```
